### 2020-04-06 README.Rmd
fixing binder

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_

<!-- badges: start -->
[![Launch Rstudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/Athanasiamo/ggsegpaper/master?urlpath=rstudio)
<!-- badges: end -->

This is the repository for the tutorial "Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_"

Full [pdf](msc_ggseg.pdf), [Rmarkdown-file](msc_ggseg.Rmd), and latex files can be found directly in this repository.

[arXiv](https://arxiv.org/abs/1912.08200)