### 2020-04-06 codethrough.Rmd
add code through

---
title: "Visualisation of Brain Statistics with R-packages _ggseg_ and _ggseg3d_"
subtite: "code through"
output: html_notebook
---

```{r setup, include=F}
knitr::opts_chunk$set(eval=TRUE, echo=TRUE, 
                      collapse = TRUE, results = 'hold', 
                      cache=FALSE, fig.pos = 'H') 
options(tinytex.verbose = TRUE, tinytex.clean = FALSE,
        tibble.width = 50) 
```

# Introduction

Visualization is increasingly important for accurate guidance and interpretation of neuroimaging results, as current research is able to generate a high amount of data and outcomes.
For Magnetic Resonance Imaging (MRI), neuroimaging software provides whole-brain information by using many small units of space (>100.000).
Nonetheless, this data is often grouped and summarized into a limited number of regions using predefined brain parcellation atlases.
Brain parcellations segment the brain into a finite set of meaningful neurobiological components, which reflect one or more brain features either based on structural or connectivity properties (@eickhoff_2018).
The use of brain atlases in neuroscience is widespread as these facilitate interpretation and minimize the amount of data (in comparison to using whole-brain data), hence reducing problems with multiple comparisons.
This enables replicability and data sharing in otherwise computationally expensive analyses, which are often performed in specialized software environments such as R (@R).

MRI data provides good spatial resolution and thus an optimal representation has to respect spatial relationships across regions.
Results from brain atlas analyses are most meaningfully visualized when projected onto a representation of the brain, thus it is desirable that any visual representation takes this relation into account.
The projection of data onto brain representations provides clear points of reference - especially when the reader is unfamiliar with the atlas - eases readability, guides interpretation, and conveys the spatial patterns of the data.
Adopting the grammar of graphics implemented in _ggplot2_ (@ggplot), one can plot neuroimaging data directly in R with several tools such as _ggBrain_ (@ggBrain) and _ggneuro_ (@ggneuro; see neuroconductor -@neuroconductor for curated neuroimaging packages for R).
Yet these tools display whole-brain image files and are not well-suited for representing brain atlas data.

In this tutorial, we introduce two packages for visualizing brain atlas data in R.
The _ggseg_ and _ggseg3d_ – plus the complimentary _ggsegExtra_ – packages, which include pre-compiled data-sets for different brain atlases that allow for 2D and 3D visualization.
The two-dimensional functionality in _ggseg_ is based on polygons and _ggplot2_-based grammar of graphics (@ggplot), while the 3D functionality in _ggseg3d_ is based on tri-surface mesh plots and _plotly_ (@plotly).

The _ggseg_ and _ggseg3d_ packages present compiled data-sets, tailored functions that allow brain data integration and plotting, and other minor features such as custom colour palettes.
The data featured in the packages are derived from two well-known parcellations: the Desikan-Killany cortical atlas (DK; @dk), which covers the cortical surface of the brain, and the Automatic Segmentation of Subcortical Structures (aseg; @aseg), which covers the subcortical structures.
Both atlases are implemented in several neuroimaging softwares, such as FreeSurfer (@fischl_99, @dale_99, @Fischl2000), and are commonly used in relation to developmental changes, disease biomarkers, genomic data, and cognition (@amlien_elaboration_2019, @WALHOVD20051261, @Pizzagalli).
The _ggsegExtra_ package contains a list of online repositories of pre-compiled atlases (currently `r sum(!grepl("None", dt$Polygon))-1` additional polygon atlases and `r sum(!grepl("None", dt$Mesh))-1` mesh atlases) and it is frequently updated.

# Tutorial
This tutorial will introduce the _ggseg_, _ggseg3d_, and _ggsegExtra_ packages and familiarize the reader with the main functions and the general use of the packages.
The tutorial will focus on the two main functions: `ggseg()` for plotting 2D polygons and `ggseg3d()` for plotting 3D brains based on tri-surface mesh plots.
The following tutorial complements the package documentation and vignettes that can be found online for all three packages: [ggseg](https://lcbc-uio.github.io/ggseg), [ggseg3d](https://lcbc-uio.github.io/ggseg3d), [ggsegExtra](https://lcbc-uio.github.io/ggsegExtra).

## Plotting polygon data (ggplot2)

`ggseg()` from the _ggseg_ package is the main function for plotting 2D data.
By default, the function automatically plots the DK atlas (see Figure \@ref(fig:figure1)).
The `ggseg()`-function is a wrapper for `geom_polygon()` from _ggplot2_, and it can be built upon and combined like any _ggplot_-object.
The image plot consists of a simple brain representation containing no extra information.
Hence, _ggseg_ plots can be easily complemented with any of the available _ggplot2_ features and options.
We recommend users to get familiarized with _ggplot2_ (@ggplot).
The package is currently only available through github, but we expect to submit the _ggseg_-package to The Comprehensive R Archive Network (@cran) in 2020.

```{r 'figure1', fig.width=3, message=FALSE, fig.cap='By default `ggseg()` will plot the dk atlas in grey shaded polygons.'}
# remotes::install_github("LCBC-UiO/ggseg")
library(ggseg)

# package combining plots easily
library(patchwork)

# Figure 1
ggseg()
```

In addition to the standard options for _ggplot2_ polygon geoms, the function also has several options for plotting the main brain representations.
These options are atlas-specific.
For cortical atlases, such as the `dk`, one can stack the hemispheres, display only the medial or lateral side, choose either one or both hemispheres, or any combination of hemisphere and view (see Figure \@ref(fig:figure2) for examples).
For subcortical atlases, such as the `aseg` atlas, the options are more limited but one can often choose between axial, sagittal, and coronal views.

(ref:figure2-cap) *ggseg* plots can be used with any *ggplot2* feature such as standard scales and themes. For cortical atlases, one can supply special *ggseg* options to determine, for example, hemisphere, view, or position. **A:** `dk` atlas, stacked with dark theme ; **B:** `dk` with medial view only; **C:** `dk` atlas with only left medial display; **D:** `dk` atlas, stacked, with classic theme; **E:** `dk` atlas with left hemisphere only; **F:** axial view of `aseg` atlas

```{r 'figure2', fig.width=6, fig.cap='(ref:figure2-cap)', out.width = "100%"}
# A: dk dark theme
fig2A <- ggseg(position = "stacked") +
  theme_dark() +
  labs(title="dk", subtitle = "dark theme")

# B: dk medial view
fig2B <- ggseg(view = "medial") +
  labs(title = "dk", subtitle = "medial view")

# C: dk left medial alone
fig2C <- ggseg(view = "medial",
            hemisphere = "left") +
  labs(title="dk", subtitle = "medial left")

# D: dk classic theme
fig2D <- ggseg(position = "stacked") +
  theme_classic() +
  labs(title = "dk", subtitle = "classic theme")

# E: dk left hemisphere
fig2E <- ggseg(hemisphere = "left") +
  labs(title = "dk", subtitle = "left hemisphere")

# F: aseg default theme
fig2F <- ggseg(atlas = aseg, view = "axial") +
  labs(title = "aseg", subtitle = "axial view")

# Using patchwork to combine all plots into one
(fig2A | fig2B | fig2C) / (fig2D | fig2E | fig2F) +
  plot_annotation(tag_levels = 'A')
```

### Using own data with fill and colour
`ggseg()` accepts any argument you can supply to `geom_polygon()`, and therefore is easy to work with for those familiar with _ggplot2_ functionality.
Standard arguments like `fill` that floods the segments with a colour, or `colour` that colours the edges around the segments are typical arguments to provide to the function either as a single setting value or within the _ggplot2_ mapping function `aes`.
To use color palettes corresponding to those used in the original neuroimaging softwares one can use atlas-specific 'brain' palette scales (Figure \@ref(fig:figure3)).

(ref:figure3-cap) Supplying 'region' to the fill option in `ggseg()`, will use the column 'region' from the accompanying data-set to create a discrete colour palette over the segments in the atlas. The `dk` atlas default palette corresponds to the FreeSurferColorLut scheme.

```{r 'figure3', fig.width=5, fig.height=5, fig.cap="(ref:figure3-cap)"}
# Figure 3
ggseg(mapping = aes(fill = region), colour="black") +
  scale_fill_brain("dk") +
  theme(legend.justification = c(1,0),
        legend.position = "bottom",
        legend.text = element_text(size = 5)) +
  guides(fill = guide_legend(ncol = 3))
```

Most users will use `ggseg()` to display - using a color scale - some descriptive or inferential statistics, such as mean thickness or brain-cognition relationships across the different brain regions.
Yet, before projecting the statistics onto the segments, one should explore the structure of the atlas data-sets.
The atlas data-set structure will help users understand what incoming statistical data needs to look like.
Note that each atlas corresponds to a unique data-set.
All data-sets have a similar structure and contain key information regarding the atlas, the region names, and the coordinates for the segment polygons.

```{r, size='small', eval=FALSE, codecap = "Inspect atlas data"}
# Look at the top 5 rows of the dk data-set
head(dk, 5)
```

In any atlas, the column 'label' is particularly useful for combining the data of interest with the _ggseg_-polygons.
The column 'label' contains the label (region) names as in the original neuroimaging software.
For example, the `dk` atlas label column matches the region names from Freesurfer statistics table outputs.
Yet, the data in _ggseg_ is in a long format - that is each region has a single row - and any data of interest needs to be in this same format.
Often data-sets are organized in wide format, in which subjects are represented by rows and each different data variable is represented in a separate column, and thus need to be rearranged to work with ggseg.
Summary statistic files from Freesurfer have a file format that can be difficult to work with in R. 
In a set of functions in _ggseg_, we facilitate conversion between FreeSurfer files and ggseg compatible data. 
There is an [extended vignette](https://lcbc-uio.github.io/ggseg/articles/freesurfer_files.html) of the various options.
In the following example, we import summary data that has been compiled using Freesurfer's `aparcstats2table` (for surfaces) or `asegstats2table` (volumes). 
The summary files can combine parcellation metrics for one or more subjects into a single summary file.
This file also contains extra metrics like intracranial volume and total grey matter volume. 
These later measures do not map onto ggseg, and as such, one should filter them out of the data returned from `read_freesurfer_table)`.

```{r, size='small'}
file_path <- "data/aparc.volume.table"

someData <- read_freesurfer_table(file_path, measure = "volume") %>% 
  # remove rows in the data that do not contain the strings "rh" or "lh"
  dplyr::filter(grepl("rh|lh", label))
someData
```

A second option is to read in atlas files directly from the entire subject directory, using regular expression to select the target atlas. 
The default aparc has a unique string ending with `aparc.stats` which we will use to read in all subject stats for that particular atlas.
One can filter specific subjects and/or regions from the R-object data.frame.

```{r, warning=FALSE}
atlas_data <- read_atlas_files(
  # Path to Freesurfers $SUBJECTS_DIR
  subjects_dir = "/Applications/freesurfer/subjects/",
  # Regular expression to find atlas files
  atlas = "aparc.stats$"
)
atlas_data
```

This long format data can be used directly with the `ggseg()`-function, as the 'label' column corresponds in name and content with the 'label' column in the atlas data of `dk`.
The data **must** include a column that has the same column name and at least _some_ data matching the values in the corresponding column in the atlas data.
In the following example, the `ggseg()`-function will recognise the matching column 'label', and merge the supplied data into the atlas using _dplyr_-joins.
The appearance of the plot can then be modified similarly to any other _ggplot2_ graph using functions such as scales, labs, themes, etc., as seen in Figure \@ref(fig:figure4)

(ref:figure4-cap) Supplying data through the '.data' option in `ggseg()` enables use of columns in the supplied data to aesthetical arguments (such as 'fill'). The *ggseg* plot can be used with any other polygon compatible function from *ggplot2* or *ggplot2* extentions, for instance adding title, changing the legend name and the colour scheme with standard *ggplot2* functions.

```{r 'figure4', fig.width=5, fig.cap="(ref:figure4-cap)", warning = FALSE}

# Figure 4A
fig4A <- someData %>% 
  # plot only subject bert
  filter(subject == "bert") %>%

  ggseg(mapping=aes(fill=volume))  +
  labs(title="from read_freesurfer_table", 
       fill="volume\n(mm^3)") +
  scale_fill_gradient(low="firebrick",high="goldenrod") +
  guides(fill = guide_colourbar(barheight = 3))

# Figure 4B
fig4B <- atlas_data %>% 
  # plot only subject bert
  filter(subject == "bert") %>%

  ggseg(mapping=aes(fill=ThickAvg))  +
  labs(title="from read_atlas_files", 
       fill="Thickness\n(mean)") +
  scale_fill_gradient(low="forestgreen",high="goldenrod")+
  guides(fill = guide_colourbar(barheight = 3))

(fig4A / fig4B) +
  plot_annotation(tag_levels = "A")
```

If the results are only in one hemisphere, but you still want to plot both of them, make sure your data.frame includes the column 'hemi' with either 'right' or 'left'.
In this case, data will be merged into the atlas both by 'region' and by 'hemi'.
For more information about adapting data and viewing only one hemisphere or side, the [online documentation](https://lcbc-uio.github.io/ggseg/articles/ggseg.html#single-hemisphere-results) contains more elaborate information.

### Creating subplots
There is often the need to plot a statistic of interest in different groups (e.g. thickness or brain - cognition relationships in young or older adults).
This may be obtained also with `ggseg()`, using _ggplot2_'s `facet_wrap` or `facet_grid`, with three guiding rules:

1. As before, data needs to be in long format with a column indexing which group the row corresponds to (group data should appear in separate rows, not in separate columns).  
2. The data needs to be grouped using _dplyr_'s `group_by()` function _before_ providing the data to the `ggseg()`-function. 
The `ggseg()`-function will detect grouped data, and adapt it to `facet`'s requirements.  
3. Apply `facet_wrap` or `facet_grid` to the plot having used the above two rules.
An example of this can be seen in Figure \@ref(fig:figure5) B, where a mock data-set including summary statistics for two groups ('Young' and 'Old') is used when faceting a _ggseg_-plot.

The second rule is needed as the `ggseg()`-function is a ggplot-wrapper rather than an extension, and thus has a particular structure and pipeline to combine the atlas to supplied data.
The result of ungrouped data into `ggseg` is shown in  \@ref(fig:figure5) A.
Note that the data is combined incorrectly, there are three panels rather than two. 
Regions not represented in the data (`NA`-values) have their own panel, while the remaining two panels only include regions with valid data.
All the concepts described above also work with the `aseg` atlas for subcortical structures, except for 'hemisphere' and 'view' arguments that are superfluous in subcortical atlases (Figure \@ref(fig:figure6)).

(ref:figure5-cap) Examples of incorrect and correct faceting with `ggseg`. **A**: ungrouped data will facet incorrectly due to the data not being merged correctly with the atlas. **B**: grouped data will merge in the expected way and result in correctly facetted plots.

```{r 'figure5', fig.width=5,  warning= FALSE, fig.cap='(ref:figure5-cap)', out.width="100%"}
# Make some simpler mock data
someData = tibble(
  region = rep(c("transverse temporal", "insula",
               "precentral","superior parietal"),2),
  p = sample(seq(0,.5,.001), 8),
  AgeG = c(rep("Young",4), rep("Old",4))
)

# Figure 5a
fig5A <- someData %>%
  ggseg(position = "stacked",
      mapping=aes(fill=p), show.legend = FALSE) +
  facet_wrap(~AgeG, ncol=3) +
  labs(title = "Ungrouped data")

# Figure 5b
fig5B <- someData %>%
  group_by(AgeG) %>%
  ggseg(position = "stacked",
      mapping=aes(fill=p), show.legend = FALSE) +
  facet_wrap(~AgeG, ncol=3) +
  labs(title = "Grouped data")

# Combining plots with patchwork, applying theme to all
(fig5A / fig5B) +
  plot_layout(widths = 1) + 
  plot_annotation(tag_levels = 'A') &
  theme(axis.text = element_text(size=8))
```

The `ggseg` package also has extensive [online documentation](https://lcbc-uio.github.io/ggseg/), with more functionality than we have covered here.
The online documentation is also regularly updated with explanations of new features or code improvements.

(ref:figure6-cap) The `aseg` atlas showing subcortical structures has some distinct differences from the `dk`. For instance, there is no option to show only a single hemisphere. Furthermore, rather than showing lateral and medial surfaces, it shows an axial and sagittal slice.

```{r 'figure6', fig.width=5, fig.cap='(ref:figure6-cap)', collapse = TRUE}
# Figure 6
ggseg(atlas="aseg", mapping=aes(fill=region)) +
  theme(legend.justification=c(1,0),
        legend.position="bottom",
        legend.text = element_text(size = 5)) +
  guides(fill = guide_legend(ncol = 3))
```

## Plotting 3D mesh data

Representing brains as 2D polygons is a good solution for fast, efficient, and flexible plotting, and can be easily combined with interactive apps such as Shiny (@shiny).
Yet, brains are intrinsically 3-dimensional and it can be challenging to recognize the location of a region as a flattened image.
This problem is exacerbated in atlases that represent subcortical features as they are 3-dimensional, while cortical structures, such as grey matter structures, can be flattened to 2-dimensions.
Hence, here we also provide the _ggseg3d_ package to plot, view, and print 3D-atlases in R.
_ggseg3d_ is based on tri-surface mesh plots using _plotly_ (@plotly).
The data structure is more complex than the _ggplot2_ polygons, and includes additional options for brain inflation, glass brains, camera locations, etc.
As _ggseg3d_ is based on plotly, the resulting brain atlases are interactive, which guides interpretation, and is useful for public dissemination.
We recommend users to familiarize themselves with _plotly_ (@plotly) when using this function.

Out-of-the-box, `ggseg3d()` plots the `dk_3d` atlas in 'LCBC' surface, but there are two more surfaces available for cortical atlases (Figure \@ref(fig:figure7))
The 'LCBC' surface consists on a semi-inflated white matter surface based on the _fsaverage5_ template subject, a surface template that is issued together with the `ggsegExtra` package.
All `[...]_3d` atlases include a `colour` column that is based on the color scheme used in the source neuroimaging software.

(ref:figure7-cap) The three surface options provided in *ggseg3d* atlases. **From left to right:** the 'white' surface is the white matter surface, 'LCBC' surface is the semi-inflated white matter surface (inflated over 10 iterations), and the 'inflated' surface is a inflated grey matter cortical surface as provided by the FreeSurfer software. The title for each plot is the literal code use to make the subplot.

```{r "figure7-show", out.width='30%',fig.show='hold',fig.ncol = 3}
# remotes::install_github("LCBC-UiO/ggseg3d")
library(ggseg3d)

# Figure 7 left
fig7l <- ggseg3d(surface = "white")  %>%
  pan_camera("right lateral") %>%
  remove_axes() %>%
  layout(title="\nggseg3d(surface='white')",
         font =  list(size = 30))
fig7l

# Figure 7 middle
fig7m <- ggseg3d()  %>%
  pan_camera("right lateral") %>%
  remove_axes() %>%
  layout(title="\nggseg3d(surface='LCBC')",
         font =  list(size = 30))
fig7m

# Figure 7 right
fig7r <- ggseg3d(surface = "inflated") %>%
  pan_camera("right lateral") %>%
  remove_axes() %>%
  layout(title="\nggseg3d(surface='inflated')",
         font =  list(size = 30))
fig7r
```

The 3D-atlas data is stored in nested tibbles.
Each cortical atlas has data-sets for three different surfaces (see Figure \@ref(fig:figure7)) and the two hemispheres.
Only one surface is available for subcortical atlases as inflation procedures are irrelevant.
The 'ggseg_3d' column includes all necessary information for `ggseg3d()` to create a 3D mesh-plot, and should not be modified by the user.
The additional 3D-atlases in _ggsegExtra_ have the same data structure.
It is important to note that the coordinates in the plot (X, Y, Z) are **not** any type of radiological coordinate system, but arbitrary Cartesian plot coordinates.
While these coordinates are important when running imaging analyses and for reporting of the placement of these regions compared to other regions, this information is not preserved in `ggseg3d`, at the moment.
As such, information regarding referencing the regions' location in standard radiological spaces (like MNI or Taliarch) should be acquired directly from imaging software or papers describing the atlases, not `ggseg3d`.

### External data supply

Similarly as in the 2D-atlas, the user will use `ggseg3d()` to display through a colour scale some descriptive or inferential statistic.
If the data is not already in the correct long format or uses similar naming as the atlas, the users should inspect the atlas data for a specific surface (and hemisphere, if desired), and then `unnest(ggseg_3d)` it to see how the atlas data is organised.

```{r out.width='80%'}
# Select surface and hemisphere, and then unnest to inspect the atlas data
dk_3d %>%
  filter(surf == "inflated" & hemi == "right") %>%
  unnest(ggseg_3d) %>%
  head(5)
```

Note the `mesh` column, which contains lists.
Each list corresponds to a region and contains 6 vectors required to create the mesh of the tri-surface plot.
It should also be noted that the 'label', 'annot' and 'region' columns could provide matching values for your own data.
Similarly to the `ggseg()`-function, the 'label' column should match the region names used in the original neuroimaging software while 'region' and 'annot' provide alternative/secondary names.
It is thus important to match your regional identifiers with those used in the atlas.
To colour the segments using a column from the data, a column name from the data needs to be supplied to the `colour` option, and providing it to the `text` option will add another line to the _plotly_ hover information.

```{r 'figure8-show', include=FALSE, out.width='40%'}
# Figure 8 left
fig8l <- ggseg3d(.data = someData, 
                 atlas = dk_3d, colour = "p", text = "p") %>% 
  pan_camera("right lateral")
fig8l

# Figure 8 middle
fig8m <- ggseg3d(.data = someData,
        atlas = dk_3d,
        colour = "p", text = "p",
        palette = c("#ff0000", "#00ff00", "#0000ff")) %>%
  pan_camera("right lateral") %>%
  remove_axes()
fig8m

# Figure 8 right
fig8r <- ggseg3d(.data = someData,
        atlas = dk_3d,
        colour = "p", text = "p",
        na.colour = "black") %>%
  pan_camera("right lateral") %>%
  remove_axes()
fig8r
```

(ref:figure8-cap) **Left:** Supplying data to `ggseg3d()` works similarly to `ggseg()`. Since `ggseg3d()` is based on *plotly* rather than *ggplot2*, most aesthetic adaptations must be set in the main function. Here we set the parcellation colours with the 'colour' column and accompanying mouse-hover text with the 'text' column. **Middle:** The palette for *ggseg3d* needs to be specified directly in the main plot call. The 'palette' option takes vectors of colours either as HEX-codes or R-colour names. **Right:** The colour of the `NA` values can also be changed through the option 'na.colour'.

### Customizing colours and the colour bar

You can provide custom colour palettes either in hex or R-names, as seen in Figure \@ref(fig:figure8).
Colours will be evenly spaced when creating the colour-scale.
A palette may also be supplied as a _named numeric vector_, where the vector names are the colours that users wish to use, and the numeric values are the breakpoints for each colour (e.g. `c("red" = 0, "white" = 0.5, "blue" = 1)`).
This way the users can control the minimum and maximum values of the colour scale, and also how the gradient is applied.
If another colour than the default gray is wanted for the `NA` regions, supply 'na.colour', either as HEX colour or colour name.
This option only takes a single colour.

### Adding a glass brain
A glass brain is a translucent representation of the brain that can provide a frame of reference,
particularly when looking at subcortical parcellations and it only serves as a visual aid.
One can control the opacity of the these `NA` structures, to improve visualization, directly in the `ggseg3d` call.
A cortical glass brain can be added to `ggseg3d` plots through the function `add_glassbrain()` (Figure \@ref(fig:figure9)), which takes three extra arguments: hemisphere, colour, and opacity.
We recommend only using the glass brain function when plotting subcortical atlases.
A similar, interactive version of this plot can be found in the [online documentation](https://lcbc-uio.github.io/ggseg3d//articles/ggseg3d.html#colours-1).

```{r 'figure9-show', out.width='40%', warning=F, message=FALSE, results = 'hide'}
# Figure 9
fig9 <- ggseg3d(atlas = aseg_3d,
        na.alpha= .5) %>%
  add_glassbrain("left") %>%
  pan_camera("left lateral")
fig9
```

(ref:figure9-cap) For subcortical structure visualization, one can add a glass brain to the plot. This will help with locating the structures relative to the cortex, and make the plot easier to interpret. The glass brain is controlled by three options: opacity, hemisphere, and colour.

```{r 'figure9', out.width='60%', echo=F, eval=T, fig.cap="(ref:figure9-cap)"}
include_plotly(fig9, "msc_ggseg_files/figure-latex/figure9.png")
```

`ggsed3d()` is based on _plotly_ and thus additional _plotly_ functionalities can be used to modify and improve the 3D atlas representations.
In addition to Carson Sievert's book on _plotly_ in R (-@plotly), we recommend resources for modifying axes in 3D plots (@plotly-ax), the basic introduction to tri-surface plots (@plotly-tri), and this tutorial on tri-surface plots with _plotly_ in R (@plotly-trisurf).
Finally, we recommend [orca](https://github.com/plotly/orca#installation) command line tool to save _ggseg3d_ atlas snapshots.

The package [online documentation](https://lcbc-uio.github.io/ggseg3d/articles/ggseg3d.html) also includes interactive versions of many of the figures in this tutorial, and several more options not covered here.
The online documentation will also be updated as new features are required or we improve the current state of user functionality and options.

## Additional atlases
The _ggseg_ and _ggseg3d_ packages have two atlases each, which are 2D and 3D variations of the same main atlases: the `dk` (@dk) and `aseg` (@aseg) atlases.
These are, however, only two among many meaningful ways of segmenting the brain into different regions.
Thus, the _ggsegExtra_ package is a library with two main functionalities: 1) to install additional data-sets for plotting with the _ggseg_ and _ggseg3d_ packages from other repositories, and 2) to create custom atlases for _ggseg_ and _ggseg3d_.

### Verified existing atlases
There is an ever-increasing amount of new atlases being created, as research and methods in neuroimaging analysis progresses.
The _ggsegExtra_-package is intended to be expanded as a community-effort, as new and informative atlases are published.
A collection of the atlases currently listed in the _ggsegExtra_ package may be found in Figure \@ref(fig:figure10) and explored in our [shinyapps.io demonstration](https://athanasiamo.shinyapps.io/ggsegDemo/).

```{r}
# remotes::install_github("LCBC-UiO/ggsegExtra")
library(ggsegExtra)

# list all known ggseg atlas-repositories
ggseg_atlas_repos()

# list repositories matching a pattern search
yeo_repo <- ggseg_atlas_repos("Yeo")
```

Having found an existing atlas, you will need to install this like an R-library.
There is a convenience-function in _ggsegExtra_ that will help you do that.
Alternatively, you can install all the existing atlases. 
This function will take time to execute and will consume quite some space as each atlas (particularly ggseg3d-atlases) can be large. 
Once the atlases are installed, the atlas-packages can be loaded like any other R-package using `library`.

```{r, eval = FALSE}
# Install ggseg-atlas listed by ggseg_atlas_repos("Yeo")
install_ggseg_atlas(yeo_repo$repo, yeo_repo$source)

# This will take quite some time, there are many atlases with heavy data.
install_ggseg_atlas_all()
```

```{r 'figure10-show', results = 'hide', message = FALSE}
library(ggsegYeo2011)
library(ggsegGlasser)
library(ggsegDesterieux)
library(ggsegICBM)

yeo7_p <- ggseg(atlas = yeo7 , mapping=aes(fill=region),
            show.legend = FALSE, position="stacked") +
  scale_fill_brain("yeo7", package = "ggsegYeo2011") +
  labs(title="yeo7-data-set")

glasser_p <- ggseg(atlas = glasser , mapping=aes(fill=region),
            show.legend = FALSE, position="stacked") +
  scale_fill_brain("glasser", package = "ggsegExtra") +
  labs(title="glasser-data-set")

ggsegs_p <- yeo7_p + glasser_p + plot_annotation(tag_levels = "A")

desterieux_p <- ggseg3d(atlas = desterieux_3d) %>%
  pan_camera("right lateral") %>%
  remove_axes() %>%
  add_annotations(
    text="desterieux_3d data-set",
    y = 1,
    yanchor = 'top',
    legendtitle=TRUE, showarrow=FALSE,
    font = list(family = 'sans serif',
                size = 50))

icbm_p <- ggseg3d(atlas = icbm_3d) %>%
  pan_camera("right lateral") %>%
  add_glassbrain("left") %>%
  remove_axes() %>%
  add_annotations(
    text="icbm_3d data-set",
    y = 1,
    yanchor = 'top',
    legendtitle=TRUE, showarrow=FALSE,
    font = list(family = 'sans serif',
                size = 50))

```

(ref:figure10-cap) Four example data-sets from the *ggsegExtra* library, plotted with the `ggseg()` and `ggseg3d()` functions.

```{r 'figure10', out.width=c('80%', '40%', '40%'), echo=FALSE, fig.show='hold', fig.ncol = 2, fig.cap='(ref:figure10-cap)'}

invisible(ggsave(filename = "msc_ggseg_files/figure-latex/figure10ab.png", 
                 plot = ggsegs_p, device = "png"))
orca(desterieux_p, file = "msc_ggseg_files/figure-latex/figure10c.png", format = "png")
orca(icbm_p, file = "msc_ggseg_files/figure-latex/figure10d.png", format = "png")

include_graphics(c(
  "msc_ggseg_files/figure-latex/figure10ab.png",
  "msc_ggseg_files/figure-latex/figure10c.png",
  "msc_ggseg_files/figure-latex/figure10d.png"
))
```

### Creating custom atlases
In many cases, users would like to create their own atlases based on existing parcellations or ones own imaging results.
There are several functions in _ggsegExtra_ to aid creating ones own atlas compatible with either _ggseg_ or _ggseg3d_. 
These processes depend on the origin of the segmentation (surface-based vs. volume-based), and what type of system you are working on. 
In addition, one will need other software dependencies to create new ggseg-compatible atlases. 
Therefore, we will not describe the process here, but refer the readers to the "Articles" sections of the [online documentation](https://lcbc-uio.github.io/ggsegExtra/index.html), for further explanations and guides. 
We encourage users who successfully create new atlases, to contribute to the _ggsegExtra_ brain atlas repository by adding the repository containing the brain atlases to the repository-list.

# Funding

This work is funded by EU Horizon 2020 Grant 'Healthy minds 0-100 years: Optimizing the use of European brain imaging cohorts (Lifebrain)', with grant agreement 732592.
The project has also received funding from the European Research Council’s Starting (grant agreements 283634, to A.M.F. and 313440 to K.B.W.) and consolidator Grant Scheme (grant agreement 771355 to KBW and 725025 to AMF).
The project has received funding through multiple grants from the Norwegian Research Council"

# Prior versions
Some of the content presented here also appears in the package vignettes of _ggseg_ and _ggseg3d_, which may be accessed through R or in the package websites (@ggseg, @ggseg3d).
Athanasia Monika Mowinckel also has several tutorials on her blog regarding _ggseg_ creation and functionality (@ggsegAnim, @ggsegIntro).

# References